### 2017-03-06 sections/implementation.tex
Update on Overleaf.

SVD decomposition is applied on matrices. The first thing to do is therefore to assemble an input matrix. Results from the Finite element method are discrete values calculated in nodes or integration points. There are usually multiple fields in result data (e.g. Displacements, Stress, Strain etc.) and each field can have multiple components (e.g. Displacements can have $X$, $Y$ and $Z$ component). As non-linear analyses usually lead to multiple computation steps there are multiple sets of data per each field component. All data corresponding to a single field component forms the input matrix to the compression algorithm. Let's call it matrix $\mtrx{A}$. Each row of matrix $\mtrx{A}$ corresponds to single time step and each column correspond to single node or integration point.

@@

Only single field component is considered to fill the input matrix $\mtrx{A}$. Results for different data components, even for the same field, can be independent from each other. There can also be order-of-magnitude difference between the components and the compression algorithm could potentially clear the important information in the weaker component. Another reason against merging of components is computational complexity. The smaller the matrix $\mtrx{A}$ is, the faster the decomposition algorithm performs.

@@

From the definition of SVD in (\ref{eq:svd-def}) and from the properties of SVD presented in section \ref{sec:math} follows the fact that a matrix can be represented in the form of its SVD components as a sum of rank 1 matrices in the form

@@

where $S_{i}$ is $i$-th singular value, $U_{i}$, $V_{i}$ are corresponding singular vectors, and $n$ is the rank of the matrix $\mtrx{A}$. Considering the fact that singular values are ordered $S_{1}>S_{2}>S_{3}> ... >S_{n}$ the above formula implies that the first term would have the highest contribution and last term would have the lowest contribution to the matrix~$\mtrx{A}$. Therefore if we take only first $r$ members of the above summation we get an approximation matrix

@@

The quality of approximation depends on the magnitude of the singular values omitted from the approximation formula, namely $S_{r+1} ...  S_{n}$. The compression algorithm is based on assumption that the first singular value is order-of-magnitude higher than singular values at the end of the decomposition sequence. In special cases, when $r=n$, or $S_{i}=0$ for all $i \geq r$ the omitted singular values do not contribute to the sum and the compression is therefore loss-less. In other cases approximation error has to be calculated and taken into account to avoid loss of important details in data.

@@

Let us assume that the matrix is not empty and has full-rank. Then from the formula above follows that if $r$ equals to the rank of matrix $\mtrx{A}$ the compression ratio is always higher than one. In other words the memory consumption of stored decomposition is bigger than the size of the original matrix. To compression algorithm be applicable the parameter $r$ must conform to the condition

@@

where $k=min(m, n)$, i.e. the smallest of two dimensions of the matrix $M$. The above formula states that the sum of squared elements of matrix $\mtrx{A}$ equals to the sum of squared singular values $\sigma_{i}$ of the same matrix $\mtrx{A}$.

Using formulas \eqref{eq:svd-expansion} and \eqref{eq:svd-approx-expansion} the equation \eqref{eq:elem-sqr-sigma-sqr} can be applied to the difference between original matrix $\mtrx{A}$ and approximation matrix $A'$.

@@

\begin{itemize}
\item The algorithm is most efficient in cases where one dimension of the input matrix is very low compared to the other. However, this is almost always the case when compressing results from FEM -- number of time steps seldom exceeds hundreds.
\item Moreover, time steps can be split into smaller ranges and the algorithm can be applied on each range separately. This will improve performance and can also increase quality of compression the key time steps on the range boundaries are carefully selected.
\item Randomized SVD algorithm has the same order of algorithmic complexity but even then can help to significantly reduce execution time.
\end{itemize}

>>>>

SVD decomposition is applied on matrices. The first thing to do is therefore to assemble an input matrix. Results from the Finite element method are discrete values calculated in nodes or integration points. There are usually multiple fields in result data (e.g. Displacements, Stress, Strain etc.) and each field can have multiple components (e.g. Displacements can have $X$, $Y$ and $Z$ component). As non-linear analyses usually lead to multiple computation steps there are multiple sets of data per each field component. All data corresponding to a single field component form the input matrix to the compression algorithm. Let's call it matrix $\mtrx{A}$. Each row of matrix $\mtrx{A}$ corresponds to single time step and each column corresponds to a single node or integration point.

@@

Only single field component is considered to fill the input matrix $\mtrx{A}$. Results for different data components (even for the same field) can be independent from each other. There can also be order-of-magnitude difference between the components and the compression algorithm could potentially clear the important information in the weaker component. Another reason against merging of components is computational complexity. The smaller the matrix $\mtrx{A}$ is, the faster the decomposition algorithm performs.

@@

From the definition of SVD in (\ref{eq:svd-def}) and from the properties of SVD presented in section \ref{sec:math} follows the fact that a matrix can be represented in the form of its SVD components as a sum of rank-1 matrices in the form

@@

where $S_{i}$ is $i$-th singular value, $\mtrx{U_{i}}$ and $\mtrx{V_{i}}$ are corresponding singular vectors, and $n$ is the rank of matrix~$\mtrx{A}$. Considering the fact that singular values are ordered $S_{1} \geq S_{2} \geq S_{3} \geq ... \geq S_{n}$ the above formula implies that the first term would have the highest contribution and last term would have the lowest contribution to matrix~$\mtrx{A}$. Therefore, if we take only first $r$ members of the above summation we get an approximation matrix

@@

The quality of approximation depends on the magnitude of the singular values omitted from the approximation formula, namely $S_{r+1} ...  S_{n}$. The compression algorithm is based on assumption that the first singular value is order-of-magnitude higher than singular values at the end of the decomposition sequence. In special cases, when $r=n$, or $S_{i}=0$ for all $i \geq r$ the omitted singular values do not contribute to the sum and the compression is therefore lossless. In other cases, approximation error has to be calculated and taken into account to avoid loss of important details in data.

@@

Let us assume that the matrix is not empty and has full-rank. Then from the formula above follows that if $r$ equals to the rank of matrix $\mtrx{A}$ the compression ratio is always higher than one. In other words the memory consumption of stored decomposition is bigger than the size of the original matrix. To make the compression algorithm applicable the parameter $r$ must conform to the condition

@@

where $k=min(m, n)$, i.e. the smallest of two dimensions of the matrix $\mtrx{M}$. The above formula states that the sum of squared elements of matrix $\mtrx{A}$ equals to the sum of squared singular values $\sigma_{i}$ of the same matrix $\mtrx{A}$.

Using formulas \eqref{eq:svd-expansion} and \eqref{eq:svd-approx-expansion} the equation \eqref{eq:elem-sqr-sigma-sqr} can be applied to the difference between original matrix $\mtrx{A}$ and approximation matrix $\mtrx{A'}$.

@@

\begin{itemize}
\item The algorithm is most efficient in cases where one dimension of the input matrix is very low compared to the other. However, this is almost always the case when compressing results from FEM -- number of time steps seldom exceeds hundreds.
\item Moreover, time steps can be split into smaller ranges and the algorithm can be applied on each range separately. This will improve performance and can also increase quality of compression if the key time steps on the range boundaries are carefully selected.
\item Randomized SVD algorithm has the same order of algorithmic complexity but yet can help to significantly reduce execution time.
\end{itemize}

### 2017-03-06 sections/math.tex
Update on Overleaf.

where $\mtrx{A}$ represents the original image and $\mtrx{A'}$ represents the
reconstructed image of dimension $m \times n$.

@@

PSNR is most commonly used to measure the quality of reconstruction of lossy compression methods (e.g., image compression). The signal in this case is the original data, and the noise is the error introduced by compression. When comparing compression methods, PSNR is an approximation to human perception of reconstruction quality. PSNR is usually expressed in terms of the logarithmic decibel scale (dB).

>>>>

where $A_{ij}$ represents an element of the original matrix and $A'_{ij}$ represents an element of the
reconstructed matrix of dimension $m \times n$.

@@

PSNR is most commonly used to measure the quality of reconstruction of lossy compression methods (e.g. image compression). The signal in this case is the original data, and the noise is the error introduced by compression. When comparing compression methods, PSNR is an approximation to human perception of reconstruction quality. PSNR is usually expressed in terms of the logarithmic decibel scale (dB).

### 2017-03-06 sections/results.tex
Update on Overleaf.

Figure \ref{fig:temelin:mesh} shows results from an analysis of aging of nuclear power-plant's containment. This analysis includes high number of analysis steps (thousands) with very little differences between them. There is therefore potential for compression ratio to be very high as proven in Figure \ref{fig:temelin:NRMSD} examining the impact of changes in the compression ratio to the mean error of approximated results.

@@

% mechaxisym
Figure \ref{fig:mechaxisym:mesh} contains visualization of results of two-dimensional analysis. There are exactly 232 analysis steps. The resulting data has linear function character with several discontinuities in temporal dimension. There are few time steps in which resulting discrete functions have very different values compared to neighboring time steps. This was supposed to have negative impact on the quality of compression. However, as can be seen in Figure \ref{fig:mechaxisym:NRMSD}, quality is better than expected. E.g. if the rank of approximation matrix is set to 3 (compared to 232 being the rank of the original matrix) the normalized relative error (NRMSD) does not exceed 1e-5.

@@

% Execution times
Besides the error also the execution speed of compression algorithm was measured. In Figure \ref{fig:temelin:ExeTime} is comparison of execution times for standard versus and SVD compression algorithms. Interestingly, execution time of standard SVD is independent of target rank whereas execution time of randomized SVD decreases linearly with decreasing target rank. If the rank is known ahead of time the fact can be taken advantage of.

>>>>

Figure \ref{fig:temelin:mesh} shows results from an analysis of aging of nuclear power-plant's containment. This analysis includes high number of analysis steps (thousands) with very little differences between them. There is therefore potential for compression ratio to be very high as proven in Figure \ref{fig:temelin:NRMSD} that examines the impact of changes in the compression ratio to the mean error of approximation.

@@

% mechaxisym
Figure \ref{fig:mechaxisym:mesh} contains visualization of results of two-dimensional analysis. There are exactly 232 analysis steps. The resulting data has linear function character with several discontinuities in temporal dimension. There are few time steps in which resulting discrete functions have very different values compared to neighboring time steps. This was supposed to have negative impact on the quality of compression. However, as can be seen in Figure \ref{fig:mechaxisym:NRMSD}, the quality is better than expected. E.g. if the rank of approximation matrix is set to 3 (compared to 232 being the rank of the original matrix) the normalized relative error (NRMSD) does not exceed 1e-5.

@@

% Execution times
Besides the error also the execution speed of compression algorithm was measured. In Figure \ref{fig:temelin:ExeTime} is comparison of execution times for standard versus randomized SVD compression algorithms. Interestingly, execution time of standard SVD is independent of target rank whereas execution time of randomized SVD decreases linearly with decreasing target rank. If the rank is known ahead of time the fact can be taken advantage of.
